% !TeX root = BTechProjectTemplate.tex
\chapter*{ACKNOWLEDGMENT}
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENT}

This project would not have taken the shape it did without the guidance of several people who gave their time and knowledge generously.

We owe our deepest thanks to our project guide, \textbf{Dr. Syed Sameen Ahmad Rizvi}, Assistant Professor, Department of Computer Science \& Engineering, whose direction shaped this work at every stage. His profound understanding of deep learning and multimodal medical imaging made sure that we remained realistic while making architectural choices, and his constructive criticism – at times positive, at times negative – always ensured that we progressed beyond what we could achieve alone.

We are truly thankful to \textbf{Dr. Priya K V}, Assistant Professor, Department of Computer Science \& Engineering, for providing us with her valuable technical support, especially in assisting us with understanding the evaluation criteria involved in super-resolution and the nuances associated with the architecture.

We thank \textbf{Dr. Murali Krishna Enduri}, Head of the Department, Computer Science \& Engineering, for ensuring that the departmental resources and research environment were available to us throughout the project.

We also thank \textbf{SRM University-AP} for giving us the opportunity to take on a project of this scope and for providing the infrastructure that made the experiments possible.

Last, and most personally, we thank our families and friends. Their encouragement on the difficult days made a real difference.

\vspace{1.5cm}
\begin{flushright}
\textbf{Santhosh Palem}\\
\textbf{Jagadeesh Ospari}\\
\textbf{Chakravarthi Padyala}
\end{flushright}