% !TeX root = BTechProjectTemplate.tex
\renewcommand{\bibname}{REFERENCES}
\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{REFERENCES}

\bibitem{srcnn} C. Dong, C. C. Loy, K. He, and X. Tang, "Image Super-Resolution Using Deep Convolutional Networks," \textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}, vol. 38, no. 2, pp. 295-307, 2016.

\bibitem{edsr} B. Lim, S. Son, H. Kim, S. Nah, and K. M. Lee, "Enhanced Deep Residual Networks for Single Image Super-Resolution," in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR) Workshops}, 2017.

\bibitem{swin_trans} Z. Liu et al., "Swin Transformer: Hierarchical Vision Transformer using Shifted Windows," in \textit{Proceedings of the IEEE/CVF International Conference on Computer Vision (ICCV)}, 2021.

\bibitem{Georgescu2023} M. Georgescu et al., "Multimodal Multi-Head Convolutional Attention With Various Kernel Sizes for Medical Image Super-Resolution," in \textit{Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision (WACV)}, 2023.

\bibitem{swinir} J. Liang, J. Cao, G. Sun, K. Zhang, L. Van Gool, and R. Timofte, "SwinIR: Image Restoration Using Swin Transformer," in \textit{Proceedings of the IEEE/CVF International Conference on Computer Vision (ICCV) Workshops}, 2021.

\bibitem{swinunetr} A. Hatamizadeh et al., "Swin UNETR: Hierarchical Vision Transformers for 3D Medical Image Segmentation," in \textit{Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision (WACV)}, 2022.

\bibitem{ixi_dataset} IXI Dataset, "Information eXtraction from Images," [Online]. Available: \url{https://brain-development.org/ixi-dataset/}.

\bibitem{lpips} R. Zhang, P. Isola, A. A. Efros, E. Shechtman, and O. Wang, "The Unreasonable Effectiveness of Deep Features as a Perceptual Metric," in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, 2018.

\bibitem{swinv2} Z. Liu et al., "Swin Transformer V2: Scaling Up Capacity and Resolution," in \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, 2022.

\bibitem{espcn} W. Shi et al., "Real-Time Single Image and Video Super-Resolution Using an Efficient Sub-Pixel Convolutional Neural Network," in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, 2016.

\bibitem{vdsr} J. Kim, J. K. Lee, and K. M. Lee, "Accurate Image Super-Resolution Using Very Deep Convolutional Networks," in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, 2016.

\bibitem{rdn} Y. Zhang, Y. Tian, Y. Kong, B. Zhong, and Y. Fu, "Residual Dense Network for Image Super-Resolution," in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, 2018.

\bibitem{idn} Z. Hui, X. Wang, and X. Gao, "Fast and Accurate Single Image Super-Resolution via Information Distillation Network," in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, 2018.

\bibitem{fscwrn} J. Shi et al., "MR Image Super-Resolution via Wide Residual Networks With Fixed Skip Connection," \textit{IEEE Journal of Biomedical and Health Informatics}, vol. 23, no. 3, pp. 1129-1140, 2019.

\bibitem{csn} X. Zhao, Y. Zhang, T. Zhang, and X. Zou, "Channel Splitting Network for Single MR Image Super-Resolution," \textit{IEEE Transactions on Image Processing}, vol. 28, no. 11, pp. 5649-5662, 2019.

\bibitem{seran} Y. Zhang, K. Li, K. Li, and Y. Fu, "MR Image Super-Resolution with Squeeze and Excitation Reasoning Attention Network," in \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, 2021.

\bibitem{t2net} C. M. Feng, H. Fu, S. Yuan, and Y. Xu, "T2-Net: A Task-to-Task Network for Joint MRI Reconstruction and Super-Resolution," in \textit{International Conference on Medical Image Computing and Computer-Assisted Intervention (MICCAI)}, 2021.

\bibitem{hat} X. Chen, X. Wang, J. Zhou, and C. Dong, "Activating More Pixels in Image Super-Resolution Transformer," in \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, 2023.

\bibitem{segsrgan} C. H. Pham et al., "Super-resolution and segmentation using generative adversarial networks—Application to neonatal brain MRI," \textit{Medical Image Analysis}, vol. 55, pp. 119-130, 2019.

\bibitem{med_sr_survey} Y. Li et al., "A Review of the Deep Learning Methods for Medical Images Super Resolution Problems," \textit{IRBM}, vol. 42, no. 2, pp. 120-133, 2021.

\end{thebibliography}
